\documentclass[../main.tex]{subfiles}

\begin{document}

\chapquote{Improving your weaknesses has the potential for the greatest gains.}{Garry Kasparov}
\chapter{Introduction}

The epigraph above -- remarked by former World Chess Champion Garry Kasparov in his most recent book \cite{kasparov2018} -- is a profound observation that applies even outside of chess.
With regard to chess in particular, Kasparov implies that you must identify your mistakes and weaknesses in order to improve as a player, and to do so, you must analyse your own games. 

Amateur chess players can analyse games they played online without much effort because the moves are recorded automatically.
However, to analyse over-the-board games\footnote{Usually, players will invest more effort in over-the-board games, both in terms of time and deep thinking. These games will also involve a greater psychological aspect as a result of being able to observe the opponent's expressions. As such, analysing these games should be even more interesting and fruitful.}, players must tediously enter the position in the computer piece by piece.
A casual over-the-board game between two friends will often reach an interesting position\footnote{For example, one of the players might have a few moves that look promising, but is also considering a line with a piece sacrifice. If he decides to play it safe, he will likely want to analyse the piece sacrifice on the computer after the game.}. 
After the game, the players will want to analyse that position on a computer, so they take a photo of the position. 
On the computer, they need to drag and drop pieces onto a virtual chessboard until the position matches the one they had on the photograph, and then they must double-check that they did not miss any pieces.

The main goal of this project is to develop a system that is able to map a photo of a chess position to a structured format that can be understood by chess engines, such as the widely-used Forsyth–Edwards Notation (FEN) \cite{edwards1994}, in order to automate this laborious task.


\input{chapters/introduction/context_survey}
% \input{chapters/introduction/objectives}

\section{Objectives}
\label{sec:objectives}
\todo{explain extent of achieving them}
\subsection{Primary}
\begin{enumerate}[label={A\arabic*.},ref={A\arabic*}]
    \item \label{obj:11} Perform a literature review of available methods for parsing chess positions from photos.
    \item \label{obj:12} Develop an algorithm for detecting the corners of the chessboard as well as the squares.
    \item \label{obj:13} Develop an algorithm for recognising the chess pieces.
    \item \label{obj:14} Develop an algorithm that uses the outputs from (2) and (3) in order to compute a probability distribution over each piece in each square.
    \item \label{obj:15} Evaluate the performance of the developed algorithms.
\end{enumerate}

\subsection{Secondary}
\begin{enumerate}[label={B\arabic*.},ref={B\arabic*}]
    \item \label{obj:21} Create a large labelled dataset of synthesised chessboard images using 3D models.
    \item \label{obj:22} Implement an algorithm that takes as input the raw probability distribution of each piece in each square and outputs a likely \gls{fen} description.
    \item \label{obj:23} Implement a simple web API that performs the inference pipeline for an input image, returning the \gls{fen} description. 
\end{enumerate}
\subsection{Tertiary}
\begin{enumerate}[label={C\arabic*.},ref={C\arabic*}]
    \item \label{obj:31} Develop a web app that allows the user to upload an image of the chess board to obtain the \gls{fen} description.
    \item \label{obj:32} Employ transfer learning to demonstrate how the system can adapt to new chess sets.
\end{enumerate}

Note: final tertiary objective was not an objective in the outstet; was added during the project, as it became clear that employing a transfer learning approach would make the system significantly more useful.


% \input{chapters/introduction/software_engineering}

\section{Software engineering process}
\todo{todo}
% A very agile approach was adopted due to the primarily research-oriented character of the project.
% Regular supervisor meetings were held where progress was discussed, and tasks were set for the next week.
% This ensured that changes could be made quickly depending on the outcomes of different approaches that were tested throughout the project.


% % first jupyter notebooks then python code for reproducibility

% Important statistics and results were persisted in data files\footnote{The data files are available in the \texttt{chesscog/results} folder.} so they could be used for analysis and plotting later.

% All code related to this project (including this report's \LaTeX{} markup itself) was maintained using the version control system Git in a single repository hosted on GitHub\footnote{The repository can be found at \href{https://github.com/georgw777/neural-surfing}{\texttt{https://github.com/georgw777/neural-surfing}}.}.
% This approach was undertaken for the reason of avoiding file and code duplication: the data files produced using the interactive notebooks could be used directly to generate the plots for the weekly notes as well as this report.
% Due to the fact that there was only one developer, all commits were made to the \texttt{master} branch; adopting a more sophisticated model utilising feature branches or pull requests would cause more overhead than benefit in the single-developer scenario.
% Furthermore, no continuous integration system with automated testing was employed because that, too, would be excessive for a research project of this scale. 

\input{chapters/introduction/ethics}

\end{document}